---
title: "Aides Drug"
author: "Shael Quintero"
date: " FAll 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```


# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

$H_0$: There is no difference in the data when comparing treatment and infection rate.

$H_A$ There is a difference in the data when comparing treatment and infection rate.



# Methods


We will use the Dyplr tool "Filter" to create a subset of the data, storing only data where drug user = "yes".
We will then use the CandP helper files to get R code to show a combination of:  graphical, numerical, and inferential to test the null hypotheseis.

```{r inlude=FALSE}
drug_users <- filter(triUmp, DrugUser == "yes")
```


<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=drug_users)
barchartGC(~infected + treatment,data=drug_users, type="percent")
```

we have created two bar charts to compare the data, the first shows total number or people who are involved in the analysis.

In chart 1 we see that most people who are drug user, do not contract Aids however it doesn't show us much about the effectiveness of the drug.

In chart 2 we have a clearer picture because it shows the data as a percentage. In this test we see that the treatment group has lower incidence of Aids compared to the placebo group.


### Numerical Descriptive Results


<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment,data=drug_users)
rowPerc(table1)
colPerc(table1)
```


In this version of the results we see the number of people in each group and the rate at which they contracted Aids.

In the infected group, we see that there are approximately three times as many cases of aids in the placebo group as opposed to the treatment group.


## Inferential Results


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion
If you discovered a relationship, be sure to take into account whether it arises because one variable causes the other or if they might be connected for some other reason.  View the GeorgeTown video's chapter 6 videos.



<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
